---
title: "Chapter 1"
author: "Paul Bertens"
date: "29-1-2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```
# Introduction
## 1
Run ggplot(data = mpg). What do you see?
```{r}
ggplot(data = mpg)
```

## 2 How may rows are in mtcars? How many columns?
```{r}
sprintf("Number of rows: %s, number of columns: %s", nrow(mtcars), ncol(mtcars))
```

## 3 What does the drv-variable describe?
drv: he type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd

## 4 Make a scatterplot of hwy versus cyl
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = hwy, y = cyl))
```

## 5
What happens if you make a scatterplot of class versus drv? Why is the plot not useful?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = class, y = drv))
```
The plot is not useful because it's a scatter of two categorical-variables data.

# Aesthetic Mappings
## 1
What's gone wrong with the code? Why are the points not blue?
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, color = "blue")
  )
```
answer
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy), color = "blue"
  )
```
## 2
Which variables in mpg are categorical? Which variables are continuous (hint: type ?mpg to read documentation for the dataset)? How can you see this information when you run mpg?
```{r}
str(mpg)
```
The character-variables and displ are categorical, hwy, cty are continuous.

## 3
Map a continuous variable to color, size and shape. How do thes aesthetics behave differently for categorical versus continuous variables.
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, color = cty)) +
  ggtitle(label = "Mapping categorical variable on aesthetic 'color'")
```

```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, size = cty)) +
  ggtitle(label = "Mapping categorical variable on aesthetic 'size'")
```

```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, shape = cty)) +
  ggtitle(label = "Mapping categorical variable on aesthetic 'shape'")
```
Continuous variables can not be mapped to a shape
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, color = manufacturer)) +
  ggtitle(label = "Mapping continuous variable on aesthetic 'color'")
```
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, size = manufacturer)) +
  ggtitle(label = "Mapping continuous variable on aesthetic 'size'")
```
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, shape = manufacturer)) +
  ggtitle(label = "Mapping continuous variable on aesthetic 'shape'")
```

## 4
What happens if you map the same variable to multiple aesthetics?
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, color = cty, size = cty)
  )
```
You get multiple legends

## 5 
What does the stroke aesthetic do? What shapes does it work with?
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy), shape = 21, color = "black", fill = "white", size = 5, stroke = 2)
```
With stroke you can change the size of the border of the shape.

## 6
What happens if you map an aesthetic to something other than a variable name, like aes(color = displ < 5)?
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, color = displ < 5)
  )
```
It returns in a logical output (TRUE when displ < 5 and FALSE when displ > 5)

# Common Problems
## 1
What happens if you facet on a continuous variable?
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy)) +
  facet_grid(rows = "cty")
```
Every value gives a facet.

## 2
What do empty cells in a plot with facet_grid(drv ~ cyl) mean? How do they relate to this plot?
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = drv, y = cyl)) +
  facet_grid(drv ~ cyl)
```
Empty cells in a facet_grid means that ther are no values in that facet. eg. there are no values in drv = r and cyl = 4

## What plots does the following code make? what does . do?
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~.)
```
It facets only in rows on variable drv (facet_grid(rows = "drv") gives the same result)

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```

It facets only in columns on variable cyl.

## 4
Take the first faceted plot in this section:
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ class, nrow = 2)
```
What are the advantages to using faceting instead of the color aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?
Advantages using facets:
- It is easier to compare the different types within a variable
- Single plot per variable
- You can compare different variables

Disadvantages using facets:
- You get a lot of single plots when there are a lot of different types within a variable

## 5
Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn't facet_grid have nrow and ncol variables?

With nrow and ncol you can set the number of rows and columns within the facet_wrap.
Other options to control the layout of individual panels:
- scales

## 6
When using facet_grid() you should usually put the variable with more unique levels in the colymns. Why?
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(manufacturer ~ cyl)
```
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(cyl ~ manufacturer)
```
The readability of plots is better.

# Geometric Objects
## 1
What geom would you use to draw a:
line chart --> geom_line()
boxplot --> geom_boxplot()
histogram --> geom_histogram()
area chart --> geom_area()

## 2
Run this code in your head and predict what the output will look like. Then run the code in R and check your predictions:
```{r}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy, col = drv)
 ) +
  geom_point() +
  geom_smooth(se = FALSE)
```

## 3
What does show.legend = FALSE do? What happens if you remove it? Why do you think it is used earlier in the chapter?
```{r}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy, color = drv)
 ) +
  geom_point() +
  geom_smooth(se = FALSE, show.legend = FALSE)
```
It removes the smoothline from the legend.
Earlier in the chapter the legend of the smoothline was removed. Main reason was that the colormapping was only used in geom_points and geom_smooth was used on all the data.

## 4
What does the se argument to geom_smooth() do?
```{r}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy)
 ) +
  geom_point(mapping = aes(color = drv)) +
  geom_smooth(se = FALSE, show.legend = FALSE) +
  ggtitle(label = "geom_smooth whit se = FALSE")
```
```{r}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy)
 ) +
  geom_point(mapping = aes(color = drv)) +
  geom_smooth(se = TRUE, show.legend = FALSE) +
  ggtitle(label = "geom_smooth whit se = TRUE")
```

## 5
Will these two graphs look different? Why/Why not?
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth()
```

```{r}
ggplot() +
  geom_point(data = mpg,
             mapping = aes(x = displ, y = hwy)) +
  geom_smooth(data = mpg,
              mapping = aes(x = displ, y = hwy))
```
There is no difference, because the data and mappings are in both situations the same. The second code is only usefull when the mappings (or the data) is different in geom_point() and geom_smooth().

## 6
Re-create the R code necessary to generate the graphs on page 21.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth(se = FALSE)
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth(se = FALSE, mapping = aes(group = drv))
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_point() +
  geom_smooth(se = FALSE)
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = drv)) +
  geom_smooth(se = FALSE)
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = drv)) +
  geom_smooth(se = FALSE, mapping = aes(linetype = drv))
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = drv)) 
```

# Statistical Transformations
## 1
What is the default geom associated with stat_summary()? How could you rewrite the R code of the plot on page 26 to use that geom function instead of the stat function?

```{r}
ggplot(data = diamonds) +
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.min = min,
    fun.max = max,
    fun = median
  )
```

The default for stat_summary() is geom_pointrange().
```{r}
ggplot(data = diamonds, mapping = aes(x = cut, y = depth)) +
  geom_pointrange(stat = "summary",
    fun.min = min,
    fun.max = max,
    fun = median)
```

## 2
What does geom_col() do? How is it different to geom_bar()? 
```{r}
ggplot(data = diamonds, mapping = aes(x = cut, y = depth)) +
  geom_col() +
  ggtitle(label = "geom_col()")
```

```{r}
ggplot(data = diamonds, mapping = aes(x = cut)) +
  geom_bar() +
  ggtitle(label = "geom_bar()")
```

geom_col() is a barchart where you can use 2 different variables.
geom_bar() is a barchart to show statistics of 1 variable. It only expects an x-variable.

## 3
Most geos and stats come in pairs that are almost always used in concert. Read through the documentation and make a list of all the pairs. What do they have in common?

The following tables lists the pairs of geoms and stats that are almost always used in concert.

| geom                       | stat                       |
|----------------------------|----------------------------|
| `geom_bar()`               | `stat_count()`             |
| `geom_bin2d()`             | `stat_bin_2d()`            |
| `geom_boxplot()`           | `stat_boxplot()`           |
| `geom_contour_filled()`    | `stat_contour_filled()`    |
| `geom_contour()`           | `stat_contour()`           |
| `geom_count()`             | `stat_sum()`               |
| `geom_density_2d()`        | `stat_density_2d()`        |
| `geom_density()`           | `stat_density()`           |
| `geom_dotplot()`           | `stat_bindot()`            |
| `geom_function()`          | `stat_function()`          |
| `geom_sf()`                | `stat_sf()`                |
| `geom_sf()`                | `stat_sf()`                |
| `geom_smooth()`            | `stat_smooth()`            |
| `geom_violin()`            | `stat_ydensity()`          |
| `geom_hex()`               | `stat_bin_hex()`           |
| `geom_qq_line()`           | `stat_qq_line()`           |
| `geom_qq()`                | `stat_qq()`                |
| `geom_quantile()`          | `stat_quantile()`          |

Table: Complementary geoms and stats

These pairs of geoms and stats tend to have their names in common, such `stat_smooth()` and `geom_smooth()` and  be documented on the same help page.
The pairs of geoms and stats that are used in concert often have each other as the default stat (for a geom) or geom (for a stat).

The following tables contain the geoms and stats in [ggplot2](https://ggplot2.tidyverse.org/reference/) and their defaults as of version 3.3.0.
Many geoms have `stat_identity()` as the default stat.


| geom                       | default stat               | shared docs |
|:---------------------------|:---------------------------|-------------|
| `geom_abline()`            | `stat_identity()`          |             |
| `geom_area()`              | `stat_identity()`          |             |
| `geom_bar()`               | `stat_count()`             | x           |
| `geom_bin2d()`             | `stat_bin_2d()`            | x           |
| `geom_blank()`             | None                       |             |
| `geom_boxplot()`           | `stat_boxplot()`           | x           |
| `geom_col()`               | `stat_identity()`          |             |
| `geom_count()`             | `stat_sum()`               | x           |
| `geom_countour_filled()`   | `stat_countour_filled()`   | x           |
| `geom_countour()`          | `stat_countour()`          | x           |
| `geom_crossbar()`          | `stat_identity()`          |             |
| `geom_curve()`             | `stat_identity()`          |             |
| `geom_density_2d_filled()` | `stat_density_2d_filled()` | x           |
| `geom_density_2d()`        | `stat_density_2d()`        | x           |
| `geom_density()`           | `stat_density()`           | x           |
| `geom_dotplot()`           | `stat_bindot()`            | x           |
| `geom_errorbar()`          | `stat_identity()`          |             |
| `geom_errorbarh()`         | `stat_identity()`          |             |
| `geom_freqpoly()`          | `stat_bin()`               | x           |
| `geom_function()`          | `stat_function()`          | x           |
| `geom_hex()`               | `stat_bin_hex()`           | x           |
| `geom_histogram()`         | `stat_bin()`               | x           |
| `geom_hline()`             | `stat_identity()`          |             |
| `geom_jitter()`            | `stat_identity()`          |             |
| `geom_label()`             | `stat_identity()`          |             |
| `geom_line()`              | `stat_identity()`          |             |
| `geom_linerange()`         | `stat_identity()`          |             |
| `geom_map()`               | `stat_identity()`          |             |
| `geom_path()`              | `stat_identity()`          |             |
| `geom_point()`             | `stat_identity()`          |             |
| `geom_pointrange()`        | `stat_identity()`          |             |
| `geom_polygon()`           | `stat_identity()`          |             |
| `geom_qq_line()`           | `stat_qq_line()`           | x           |
| `geom_qq()`                | `stat_qq()`                | x           |
| `geom_quantile()`          | `stat_quantile()`          | x           |
| `geom_raster()`            | `stat_identity()`          |             | 
| `geom_rect()`              | `stat_identity()`          |             |  
| `geom_ribbon()`            | `stat_identity()`          |             |
| `geom_rug()`               | `stat_identity()`          |             |
| `geom_segment()`           | `stat_identity()`          |             |
| `geom_sf_label()`          | `stat_sf_coordinates()`    | x           |
| `geom_sf_text()`           | `stat_sf_coordinates()`    | x           |
| `geom_sf()`                | `stat_sf()`                | x           |
| `geom_smooth()`            | `stat_smooth()`            | x           |
| `geom_spoke()`             | `stat_identity()`          |             |
| `geom_step()`              | `stat_identity()`          |             |
| `geom_text()`              | `stat_identity()`          |             |
| `geom_tile()`              | `stat_identity()`          |             |  
| `geom_violin()`            | `stat_ydensity()`          | x           |
| `geom_vline()`             | `stat_identity()`          |             |

Table: ggplot2 geom layers and their default stats.

| stat                        | default geom               | shared docs |
|:----------------------------|:---------------------------|-------------|
| `stat_bin_2d()`             | `geom_tile()`              |             |
| `stat_bin_hex()`            | `geom_hex()`               | x           |
| `stat_bin()`                | `geom_bar()`               | x           |
| `stat_boxplot()`            | `geom_boxplot()`           | x           |
| `stat_count()`              | `geom_bar()`               | x           |
| `stat_countour_filled()`    | `geom_contour_filled()`    | x           |
| `stat_countour()`           | `geom_contour()`           | x           |
| `stat_density_2d_filled()`  | `geom_density_2d()`        | x           |
| `stat_density_2d()`         | `geom_density_2d()`        | x           |
| `stat_density()`            | `geom_area()`              |             |
| `stat_ecdf()`               | `geom_step()`              |             |
| `stat_ellipse()`            | `geom_path()`              |             |
| `stat_function()`           | `geom_function()`          | x           |
| `stat_function()`           | `geom_path()`              |             |
| `stat_identity()`           | `geom_point()`             |             |
| `stat_qq_line()`            | `geom_path()`              |             |
| `stat_qq()`                 | `geom_point()`             |             |
| `stat_quantile()`           | `geom_quantile()`          | x           |
| `stat_sf_coordinates()`     | `geom_point()`             |             |
| `stat_sf()`                 | `geom_rect()`              |             |
| `stat_smooth()`             | `geom_smooth()`            | x           |
| `stat_sum()`                | `geom_point()`             |             |
| `stat_summary_2d()`         | `geom_tile()`              |             |
| `stat_summary_bin()`        | `geom_pointrange()`        |             |
| `stat_summary_hex()`        | `geom_hex()`               |             |
| `stat_summary()`            | `geom_pointrange()`        |             |
| `stat_unique()`             | `geom_point()`             |             |
| `stat_ydensity()`           | `geom_violin()`            | x           |

Table: ggplot2 stat layers and their default geoms.

## 4
What variable does `stat_smooth()` compute? What parameters control it's behavior?
This geom treats each axis differently and, thus, can thus have two orientations. Often the orientation is easy to deduce from a combination of the given mappings and the types of positional scales in use. Thus, ggplot2 will by default try to guess which orientation the layer should have. Under rare circumstances, the orientation is ambiguous and guessing may fail. In that case the orientation can be specified directly using the orientation parameter, which can be either "x" or "y". The value gives the axis that the geom should run along, "x" being the default orientation you would expect for the geom.

method	
Smoothing method (function) to use, accepts either NULL or a character vector, e.g. "lm", "glm", "gam", "loess" or a function, e.g. MASS::rlm or mgcv::gam, stats::lm, or stats::loess. "auto" is also accepted for backwards compatibility. It is equivalent to NULL.

For method = NULL the smoothing method is chosen based on the size of the largest group (across all panels). stats::loess() is used for less than 1,000 observations; otherwise mgcv::gam() is used with formula = y ~ s(x, bs = "cs") with method = "REML". Somewhat anecdotally, loess gives a better appearance, but is O(N^2) in memory, so does not work for larger datasets.

If you have fewer than 1,000 observations but want to use the same gam() model that method = NULL would use, then set method = "gam", formula = y ~ s(x, bs = "cs").

formula	
Formula to use in smoothing function, eg. y ~ x, y ~ poly(x, 2), y ~ log(x). NULL by default, in which case method = NULL implies formula = y ~ x when there are fewer than 1,000 observations and formula = y ~ s(x, bs = "cs") otherwise.

se	
Display confidence interval around smooth? (TRUE by default, see level to control.)

na.rm	
If FALSE, the default, missing values are removed with a warning. If TRUE, missing values are silently removed.

orientation	
The orientation of the layer. The default (NA) automatically determines the orientation from the aesthetic mapping. In the rare event that this fails it can be given explicitly by setting orientation to either "x" or "y". See the Orientation section for more detail.

## 5
In our proportion bar chart, we need to set group = 1. Why? In other words what is the problem with these two graphs?

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, y = ..prop..)) +
  ggtitle(label = "without group = 1")
```
If `group = 1` is not included, then all the bars in the plot will have the same height, a height of 1.
The function `geom_bar()` assumes that the groups are equal to the `x` values since the stat computes the counts within the group.

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1)) +
  ggtitle(label = "with group = 1")
```


```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop..)) +
  ggtitle(label = "without group = 1")
```
With the `fill` aesthetic, the heights of the bars need to be normalized.
```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = color, y = ..count.. / sum(..count..))) +
  ggtitle(label = "with normalized bars")
```


