---
title: "Chapter 1"
author: "Paul Bertens"
date: "29-1-2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```
# Introduction
## 1
Run ggplot(data = mpg). What do you see?
```{r}
ggplot(data = mpg)
```

## 2 How may rows are in mtcars? How many columns?
```{r}
sprintf("Number of rows: %s, number of columns: %s", nrow(mtcars), ncol(mtcars))
```

## 3 What does the drv-variable describe?
drv: he type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd

## 4 Make a scatterplot of hwy versus cyl
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = hwy, y = cyl))
```

## 5
What happens if you make a scatterplot of class versus drv? Why is the plot not useful?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = class, y = drv))
```
The plot is not useful because it's a scatter of two categorical-variables data.

# Aesthetic Mappings
## 1
What's gone wrong with the code? Why are the points not blue?
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, color = "blue")
  )
```
answer
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy), color = "blue"
  )
```
## 2
Which variables in mpg are categorical? Which variables are continuous (hint: type ?mpg to read documentation for the dataset)? How can you see this information when you run mpg?
```{r}
str(mpg)
```
The character-variables and displ are categorical, hwy, cty are continuous.

## 3
Map a continuous variable to color, size and shape. How do thes aesthetics behave differently for categorical versus continuous variables.
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, color = cty)) +
  ggtitle(label = "Mapping categorical variable on aesthetic 'color'")
```

```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, size = cty)) +
  ggtitle(label = "Mapping categorical variable on aesthetic 'size'")
```

```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, shape = cty)) +
  ggtitle(label = "Mapping categorical variable on aesthetic 'shape'")
```
Continuous variables can not be mapped to a shape
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, color = manufacturer)) +
  ggtitle(label = "Mapping continuous variable on aesthetic 'color'")
```
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, size = manufacturer)) +
  ggtitle(label = "Mapping continuous variable on aesthetic 'size'")
```
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, shape = manufacturer)) +
  ggtitle(label = "Mapping continuous variable on aesthetic 'shape'")
```

## 4
What happens if you map the same variable to multiple aesthetics?
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, color = cty, size = cty)
  )
```
You get multiple legends

## 5 
What does the stroke aesthetic do? What shapes does it work with?
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, shape = factor(year), stroke = 1)
  )
```
With stroke you can change the size of the shape.

## 6
What happens if you map an aesthetic to something other than a variable name, like aes(color = displ < 5)?
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, color = displ < 5)
  )
```
It returns in a logical output (TRUE when displ < 5 and FALSE when displ > 5)

# Common Problems
## 1
What happens if you facet on a continuous variable?
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy)) +
  facet_grid(rows = "cty")
```
Every value gives a facet.

## 2
What do empty cells in a plot with facet_grid(drv ~ cyl) mean? How do they relate to this plot?
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = drv, y = cyl)) +
  facet_grid(drv ~ cyl)
```
Empty cells in a facet_grid means that ther are no values in that facet. eg. there are no values in drv = r and cyl = 4

## What plots does the following code make? what does . do?
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~.)
```
It facets only in rows on variable drv (facet_grid(rows = "drv") gives the same result)

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```

It facets only in columns on variable cyl.

## 4
Take the first faceted plot in this section:
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ class, nrow = 2)
```
What are the advantages to using faceting instead of the color aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?
Advantages using facets:
- It is easier to compare the different types within a variable
- Single plot per variable
- You can compare different variables

Disadvantages using facets:
- You get a lot of single plots when there are a lot of different types within a variable

## 5
Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn't facet_grid have nrow and ncol variables?

With nrow and ncol you can set the number of rows and columns within the facet_wrap.
Other options to control the layout of individual panels:
- scales

## 6
When using facet_grid() you should usually put the variable with more unique levels in the colymns. Why?
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(manufacturer ~ cyl)
```
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(cyl ~ manufacturer)
```
The readability of plots is better.

