---
title: "Chapter 1"
author: "Paul Bertens"
date: "29-1-2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```
# Exercises page 6 + 7
## 1
Run ggplot(data = mpg). What do you see?
```{r}
ggplot(data = mpg)
```

## 2 How may rows are in mtcars? How many columns?
```{r}
sprintf("Number of rows: %s, number of columns: %s", nrow(mtcars), ncol(mtcars))
```

## 3 What does the drv-variable describe?
drv: he type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd

## 4 Make a scatterplot of hwy versus cyl
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = hwy, y = cyl))
```

## 5
What happens if you make a scatterplot of class versus drv? Why is the plot not useful?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = class, y = drv))
```
The plot is not useful because it's a scatter of two categorical-variables data.

# Exercises page 12 + 13
## 2
What's gone wrong with the code? Why are the points not blue?
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, color = "blue")
  )
```
answer
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy), color = "blue"
  )
```
## 2
Which variables in mpg are categorical? Which variables are continuous (hint: type ?mpg to read documentation for the dataset)? How can you see this information when you run mpg?
```{r}
str(mpg)
```
The character-variables and displ are categorical, hwy, cty are continuous.

## 3
Map a continuous variable to color, size and shape. How do thes aesthetics behave differently for categorical versus continuous variables.
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, color = cty)) +
  ggtitle(label = "Mapping categorical variable on aesthetic 'color'")
```

```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, size = cty)) +
  ggtitle(label = "Mapping categorical variable on aesthetic 'size'")
```

```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, shape = cty)) +
  ggtitle(label = "Mapping categorical variable on aesthetic 'shape'")
```
Continuous variables can not be mapped to a shape
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, color = manufacturer)) +
  ggtitle(label = "Mapping continuous variable on aesthetic 'color'")
```
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, size = manufacturer)) +
  ggtitle(label = "Mapping continuous variable on aesthetic 'size'")
```
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, shape = manufacturer)) +
  ggtitle(label = "Mapping continuous variable on aesthetic 'shape'")
```

## 4
What happens if you map the same variable to multiple aesthetics?
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, color = cty, size = cty)
  )
```
You get multiple legends

## 5 
What does the stroke aesthetic do? What shapes does it work with?
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, shape = factor(year), stroke = 1)
  )
```
With stroke you can change the size of the shape.

## 6
What happens if you map an aesthetic to something other than a variable name, like aes(color = displ < 5)?
```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, color = displ < 5)
  )
```
It returns in a logical output (TRUE when displ < 5 and FALSE when displ > 5)



